---
title: "GIS Lab 10"
author: "Camille Okonkwo"
output:
  pdf_document:
    latex_engine: xelatex
    toc: yes
    toc_depth: 2
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '2'
header-includes:
- \usepackage{fancyhdr}
- \usepackage{lipsum}
- \pagestyle{fancy}
- \fancyhead[R]{\thepage}
- \fancypagestyle{plain}{\pagestyle{fancy}}
- \usepackage{fontspec}
editor_options: 
  chunk_output_type: console
--- 
\newpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Packages 
```{r message=FALSE, warnings=FALSE}
library(sf)
library(tmap)
library(tidyverse)
library(spgwr)
library(jtools)
library(tmaptools)
library(broom.mixed)
```

# Read in shapefile
```{r}
data = st_read ("data/NYC_zip_COVID.shp")
```
\newpage

# Scatter plots 
```{r}
# Essential Workers
data |> 
  ggplot(aes(x = per_ssn,  y = deth_rt, color = CovidZip_2)) +
  geom_point() + 
  labs(title = "Percent Essential Worker and COVID-19 Death Rate",
       x = "Percent Essential Worker",
       y = "Death Rate",
       color = "Borough") +
  theme_minimal()

# Add linear fit and facet
data  |>  
  ggplot(aes(x = per_ssn,  y = deth_rt, color = CovidZip_2)) +
  geom_point() + 
  geom_smooth(method = lm, se = FALSE) +
  labs(title = "Percent Essential Worker and COVID-19 Death Rate",
       x = "Percent Essential Worker",
       y = "Death Rate",
       color = "Borough") +
  facet_wrap(~CovidZip_2)  +
  theme_minimal() +
  theme(legend.position="bottom")

# Percent 65yrs and above 
data |> 
  ggplot(aes(x = pr_pp65,  y = deth_rt, color = CovidZip_2 )) +
  geom_point() + 
  labs(title = "Percent over 65yrs and COVID-19 Death Rate",
       x = "Percent over 65yrs",
       y = "Death Rate",
       color = "Borough") +
  theme_minimal()


# Add linear fit and facet
data |> 
  ggplot(aes(x = pr_pp65,  y = deth_rt, color = CovidZip_2 )) +
  geom_point() + 
  geom_smooth(method = lm, se = FALSE) +
  labs(title = "Percent over 65yrs and COVID-19 Death Rate",
       x = "Percent over 65yrs",
       y = "Death Rate",
       color = "Borough") +
  facet_wrap(~CovidZip_2)  +
  theme_minimal() +
  theme(legend.position="bottom")
```
\newpage

# Multiple Regression
```{r}
mreg = lm(deth_rt ~ pr_pp65 + per_ssn + pr_hghd + med_inc  + pr_pbtr,
           data = data) 


summary(mreg)
AIC(mreg)


plot_summs(mreg, scale = TRUE, inner_ci_level = .95)
```
\newpage

# Geographically-Weighted Regression 
```{r}
# Fixed Bandwidth 
class(data)
data = as_Spatial(data)


fixed_bw = gwr.sel(deth_rt ~ pr_pp65 + per_ssn + pr_hghd + med_inc  + pr_pbtr,
               data = data, 
               adapt=FALSE, 
               gweight = gwr.Gauss,
               verbose = TRUE)


fbw_gwr = gwr(deth_rt ~ pr_pp65 + per_ssn + pr_hghd + med_inc  + pr_pbtr,
               data = data, 
               bandwidth = fixed_bw,
               gweight=gwr.Gauss,
               hatmatrix=TRUE, 
               se.fit=TRUE)

fbw_gwr

names(fbw_gwr)

# Grab results 
results = fbw_gwr$SDF

names(results)
summary(results)
```
\newpage
# Mapping Results

## Local R-Square
```{r}
# Local R Square
r2 = 
  tm_shape(results) +
  tm_polygons(col = "localR2",
              palette = "magma",
              style = "cont",
              border.col = "black",
              lwd = .5,
              title = "R-Square") +
  tm_legend(legend.position = c(0.05, 0.450),
            legend.text.size = 1) +
  tm_layout(title = "Local R-Square",
            title.size = 2,                                               
            title.position = c(0.0, .89),
            frame = FALSE) +
  tm_credits("Author: Camille Okonkwo \nSource: NYC Health \nDate: 04/10/2024",
             size = .7,
             position = c(.66,.04))

r2

tmap_save(r2, 
          dpi = 300,
          width = 6,
          height = 6,
          filename = "local_r2.png")
```
\newpage
## Percent 65 years and over 
```{r}
coef_65yrs =
  tm_shape(results) +
  tm_polygons(col = "pr_pp65",
              palette = "magma",
              style = "cont",
              border.col = "black",
              lwd = .5,
              title = "Coefficient") +
  tm_legend(legend.position = c(0.18, 0.50),
            legend.text.size = 1) +
  tm_layout(title = "Percent 65 \nand Over",
            title.size = 2,                                                  
            title.position = c(0.12, .85),  
            frame = FALSE) +
  tm_credits("Author: Camille Okonkwo \nSource: NYC Health \nDate: 04/10/2024",
             size = .7,
             position = c(.66,.04))

coef_65yrs

tmap_save(coef_65yrs, 
          dpi = 300,
          width = 6,
          height = 6,
          filename = "65yr_coef.png")
```

### Significance of 65 years and over 
```{r}
results$tval_65 = results$pr_pp65/results$pr_pp65_se

results$t_65_cat = cut(results$tval_65,
                             breaks=c(min(results$tval_65), -2, 2,
                                      max(results$tval_65)),
                             labels=c("sig","nonsig", "sig"))

tm_shape(results) + 
  tm_fill(col = "t_65_cat", 
          title = "Statistical Significant",
          legend.hist = TRUE) 
```
\newpage
## Percent Essential Workers
```{r}
essen = tm_shape(results) +
  tm_polygons(col = "per_ssn",
              palette = "magma",
              style = "cont",
              border.col = "black",
              lwd = .5,
              title = "Coefficient") +
  tm_legend(legend.position = c(0.18, 0.50),
            legend.text.size = 1) +
  tm_layout(title = "Percent Essential \nWorkers",
            title.size = 2,                                             
            title.position = c(0.0, .9),  
            frame = FALSE) +
  tm_credits("Author: Camille Okonkwo \nSource: NYC Health \nDate: 04/10/2024",
             size = .7,
             position = c(.66,.04))

essen

tmap_save(essen, 
          dpi = 300,
          width = 6,
          height = 6,
          filename = "essential_coef.png")
```

### Significance of Essential Workers
```{r}
results$tval_essen = results$per_ssn/results$per_ssn_se

results$t_essen_cat = cut(results$tval_essen,
                        breaks=c(min(results$tval_essen), -2, 2, max(results$tval_essen)),
                        labels=c("sig","nonsig", "sig"))

tm_shape(results) + 
  tm_fill(col = "t_essen_cat", 
          title = "Statistical Significant",
          legend.hist = TRUE) 
```

## Percent High School Diploma
```{r}
hs_diploma = tm_shape(results) +
  tm_polygons(col = "pr_hghd",
              palette = "magma",
              style = "cont",
              midpoint = 0,
              border.col = "black",
              lwd = .5,
              title = "Coefficient") +
  tm_legend(legend.position = c(0.0, 0.45),
            legend.text.size = 1) +
  tm_layout(title = "Percent HS \nDiploma",
            title.size = 2,                                             
            title.position = c(0., .85),  
            frame = FALSE) +
  tm_credits("Author: Camille Okonkwo \nSource: NYC Health \nDate: 04/10/2024",
             size = .7,
             position = c(.66,.04))

hs_diploma

tmap_save(hs_diploma, 
          dpi = 300,
          width = 6,
          height = 6,
          filename = "hs_diploma.png")
```

### Significance of High School Diploma
```{r}
results$tval_hs = results$pr_hghd/results$pr_hghd_se

results$t_hs_cat = cut(results$tval_hs,
                           breaks=c(min(results$tval_hs), -2, 2, max(results$tval_hs)),
                           labels=c("sig","nonsig", "sig"))

tm_shape(results) + 
  tm_fill(col = "t_hs_cat", 
          title = "Statistical Significant",
          legend.hist = TRUE) 
```

